---
title: "Bellabeats_Case_Study"
author: "Stephanie Sun"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Phases of Data Analysis**

### Bellabeats Case Study for Google Data Analytics Certificate Program

## 1. The Ask

###### -Business Task: How can the client, Bellabeats, grow their market share in the Smart Device industry?

###### -Insight into competitor, FitBit, consumer behaviour will inform opportunities for growth by finding patterns and understanding needs.

###### -Stakeholders: Bellabeat founders - Urska and Sando

## 2. Prepare The Data

##### -Client suggests FitBit data from Kaggle by Mobius

##### -Additional Google Trends data can be used to inform marketing and advertising decisions

##### -Stored in Google Drive Folder for easy access with Sheets in order to verify data integrity

##### -Limitations:Sample size is small and is not a strong representation of the population

## 3. Process The Data

###### -Spreadsheets used to clean data (looked for nulls, inconsistencies, incomplete columns)

###### -R will be used for data exploration

###### -Cleaning documented in notes tab of csv files

## 4. Data Analysis 

###### -Cleaned datasets downloaded as csv files and imported into R Studio
###### -Analysis completed in R Studio desktop
###### -**Most users are a healthy BMI and they tracked their daily activity but not weight or sleep**

## 5. Share Findings

###### -Continue scrolling down in this markdown to view findings

## 6. Recommendations and Next Steps

###### -Bellabeats can grow their market share by increasing their advertising campaigns during the holidays to take advantage of increased interest for their competitor during that time
###### -Messaging can be directed towards healthy users, daily activity tracking, or evening/noon workouts
###### -Opportunity to grow to new users by researching why there are less overweight users and sleep/weight tracking
###### -A large sample size and greater timeframe for data collected would provide stronger representation of the population

# **Data Exploration, Visualizations, and Findings**

```{r message=FALSE, warning=FALSE, include=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

```{r install and load packages, message=FALSE, warning=FALSE}

install.packages("tidyverse")

library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
```

```{r import datasets, message=FALSE, warning=FALSE}
fit_bit_gtrend <- read_csv('FitBit google trend - cleaned.csv')

activity <- read_csv('y dailyActivity_merged - dailyActivity_merged.csv')

sleep <- read_csv('y sleepDay_merged - cleaned.csv')

weight <- read_csv('y weightLogInfo_merged - cleaned.csv')

hourly_steps <- read.csv('y hourlySteps_merged - cleaned.csv')

steps <- read.csv('hourly_steps.csv')
```

#### Only 1 record in the overweight BMI category

```{r Joining Activity and Weight dataframes, warning=FALSE}

activity_weight <- inner_join(activity, weight, by = "Id") %>% 
  select(Id, TotalSteps, SedentaryMinutes, BMI, WeightPounds) %>% 
  group_by(Id) %>% 
  summarize(mean_steps=mean(TotalSteps,na.rm = T),
            mean_sedentary_minutes=mean(SedentaryMinutes,na.rm = T),
            mean_bmi=mean(BMI,na.rm = T),
            mean_weightlbs=mean(WeightPounds,na.rm = T))  

activity_weight

```

#### -Not enough user data to determine if correlation exists between weight and steps

```{r Plot Compares User Weight and Average Daily Steps, echo=FALSE}

ggplot(data=activity_weight, mapping = aes(x=mean_steps, y=mean_weightlbs))+
  geom_point() +
  labs(title="User Weight and Average Daily Steps",
       y="User Weight in lbs", x="Avg. Daily Steps")

```

#### -Dataframe, daily_activity, was created to aggregate average steps by date

```{r Summary table of Average Daily Steps}

daily_activity <- activity %>% 
  group_by(ActivityDate) %>%
  summarize(mean_steps=mean(TotalSteps))

daily_activity

```

#### -Sorted daily_activity dataframe by steps to determine if there were more or less activity on certain days of the week

#### -Daily Activity data also visualized in a graph

#### -No correlations or patterns found, data is also insufficient

```{r daily activities sorted by steps}

daily_activity %>% 
  arrange(mean_steps)

```

```{r visualization of daily activities by date, echo=FALSE}

ggplot(data=daily_activity, aes(x=ActivityDate, y=mean_steps)) +
  geom_bar(stat = "identity")+
  labs(title="Average Steps by Date", y="Avg. Steps", x="Date")+
  theme(axis.text.x = element_text(angle = 90))

```

#### -Next hourly activity was table and graph were created to see if when users were more active

#### -Found that activity was highest in the evenings and around noon with 6pm being the most active hour

```{r determine hour of day with most steps}

summary_steps <- steps %>% 
  group_by(Time) %>%
  summarize(mean_steps=mean(StepTotal)) %>% 
  arrange(Time)

summary_steps

```

```{r chart of hourly activity, echo=FALSE}

ggplot(data=summary_steps, aes(x=Time, y=mean_steps)) +
  geom_bar(stat = "identity")+
  labs(title="Average Steps by Hour of Day", y="Avg. Steps", x="Time (24hr format)")

```

#### -Dataframes were created and joined to compare the total logs tracked

#### -Pivoted to a longer table to visualize in a bar chart

#### -Most users tracked their activity, some tracked their sleep, and few tracked their weight

#### -*data tracked over a month, count of 31 indicated daily tracking*

```{r comparing the count of activity, sleep, weight}

activity_logs<- activity %>% #most users track daily
  count(Id)

sleep_logs<- sleep %>% #some users track
  count(Id)


weight_logs<- weight %>% #few users track
  count(Id)


logs <- full_join(activity_logs, sleep_logs, by = "Id") %>%
  full_join(weight_logs, by = "Id") %>% 
  rename(activity = n.x, sleep = n.y, weight=n)

logs

```

```{r long chart format and graph to compare count of logs by type, echo=FALSE, warning=FALSE}

logs_long <- logs %>%
  pivot_longer(cols = c(activity, sleep, weight), names_to = "type", values_to = "count")

logs_bar<- ggplot(logs_long, aes(x = type, y = count, fill = type)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Logs Recorded by Type", x = "Type", y = "Total Logs")

logs_bar

```

#### -Google Trend graph plotted to show when interest in FitBit is at its peak (December/holidays)

#### -This may be something Bellabeats marketing and advertising teams may want to take advantage of

```{r line graph of FitBit interest since 2013, echo=FALSE}

ggplot(data=fit_bit_gtrend, mapping = aes(x=Week, y=interest_over_time))+
  geom_line()+
  labs(title="FitBit Google Trends Data", y="Interest Over Time")

```
