---
title: "Dogs & Intelligence"
author: "Stephanie Sun"
date: "2023-02-23"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## Install and load packages

```{r read csv, message=FALSE}

install.packages("tidyverse")
library(readr) 
library(ggplot2)
library(dplyr)

```

## SQL queried csv file imported


```{r read csv and view the dataframe, message=FALSE, warning=FALSE}

dogs <- read_csv("dog_intelligence_size_v3_R.csv")

```
## Analysis of Dataframe

##### -started creating a dataframe consisting of average weight by classifitaion 
##### which was used to create a bar graph to visualize average weight vs intelligence classification


```{r dataframe with average weight by classification, message=FALSE, warning=FALSE}

class_weight <- dogs %>% 
  group_by(classification) %>%
  summarize(avg_weight=mean(median_weight)) %>% 
  arrange(desc(avg_weight))

class_weight

```

```{r bar graph showing average weight of each classification, message=FALSE, warning=FALSE}

dog_barchart <- ggplot(data = class_weight) +
  geom_bar(mapping = aes(y = reorder(classification, avg_weight), x = avg_weight, 
                         fill=classification), stat = "identity")+
  labs(title="Weight of Classification", y="Classification", x="Average Weight lbs")+
  guides(fill=FALSE)

dog_barchart

```


##### -scatter plots created to illustrate relationship between obedience and repetitions,
##### included classification in the plot because it groups dogs based on obedience and repetitions


```{r dog intelligence scatterplot, message=FALSE, warning=FALSE}

dog_intelligence <- ggplot(data=dogs)+
  geom_point(mapping = aes(x = obey_rate, y = median_reps,
                           color = classification, size = 0.6))+
  facet_wrap(~obey_rate~median_reps)+
  labs(title="Obey vs Reps", subtitle="By Classification",
       x="Obey On First Command %",
       y="Median Repetitions to Learn New Command")+
  guides(size=FALSE)

dog_intelligence

```

##### -Scatter plot answers the question and illustrates the correlation of dog size vs intelligence



```{r dog size and intelligence scatterplot, message=FALSE, warning=FALSE}

dog_size <- ggplot(data=dogs)+
  geom_point(mapping=aes(x=median_height, y=median_weight,
                         color=classification))+
  geom_smooth(mapping=aes(x=median_height, y=median_weight),
              color="black", size=.5)+
  labs(title="Dog Size and Intelligence Classification",
       x="Median Weight (lbs)", y="Median Height (in)")

dog_size

```




